Blockchain System

### 2020-12-14 papers/execution-model/model.tex
final-touchup

\begin{definition}%[Program]
\emph{Expressions} are defined by the following grammar:
\begin{align*}
  \CONSTANT & ::= \ADDR \mid \OPH \mid \PUH \mid\PUK \mid \CODE \mid
              \NTEZ \mid \SUNIT \mid \FALSE \mid \TRUE \\
  \STATUS & ::= \STATUSPENDING \mid \STATUSINCLUDING (\dots) \mid \STATUSTIMEOUT \\
	\EXPR & :: =  \CONSTANT \mid \STATUS \mid \VARIABLE \mid \lambda \VARIABLE. \EXPR\ 
	\mid \EXPR\EXPR \mid (\EXPR,\EXPR) \mid \FST\ \EXPR \mid \SND\ \EXPR 	%\mid \NI \EI
	\\&\mid \EXPR\ \PLUS\ \EXPR\ \mid \EXPR\ \MINUS\ \EXPR\ \mid \EXPR\ \EQUAL\ \EXPR\ \mid \EXPR\ \LESS\ \EXPR
	\\&\mid \EXPR\ \AND\ \EXPR\ \mid \EXPR\ \OR\ \EXPR\ \mid \NOT\ \EXPR
  \\&\mid \NIL\ \mid \CONS\ \EXPR\ \EXPR \mid \LEFT\ \EXPR \mid\RIGHT\
  \EXPR \mid \SOME\ \EXPR \mid \NONE
  \\& \mid \RAISE\ \EXPR \mid \TRY\ \EXPR\ \EXCEPT\ \EXPR
  \\& \mid \MATCH\ \EXPR\ \WITH\ \PATTERN\to\EXPR \dots
  \\& \mid \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR
  \\& \mid \ORIGINATE\EXPR\EXPR\EXPR\EXPR\EXPR
  \\&\mid \QOP\  \EXPR\\
  \QOP & ::= balance \mid status \mid storage \mid code \mid counter
\end{align*}
\end{definition}

@@

\begin{definition}%[Blockchain system]
A \emph{blockchain configuration} is a pair
$ \NODE \| \BLOCKCHAIN$ where
\fxnote[inline]{Should the model contain multiple nodes?}
\begin{enumerate}
\item $\NODE = [\ACCOUNTS, \OPERATIONS]$ is the state of a node, and
\item $\BLOCKCHAIN = [\PENDING, \ACCEPTED, \MANAGERS, \CONTRACTORS, \TIME]$ is the state of a blockchain 
  such that $\forall c \in \ACCOUNTS \implies \exists k\in \MANAGERS, k.\PUK = c.\PUK$.
\end{enumerate}
\end{definition}

@@

\begin{itemize}
\item library call \texttt{call\_contract} with \texttt{confirmations:None} corresponds to
  \begin{enumerate}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Reject(balance,count,pubkey,fee)}
  \end{enumerate}
\item afterwards must check the status to find whether the
  transfer is still pending or \TirName{Block-Accept''} or
  \TirName{Block-Timeout''} has happened.
\item no further transfer is possible before the status is resolved to
  accept or timeout.
\end{itemize}

@@

    \NodeTrans\\ [\epsilon[\FAIL] ::\EXPRS, \ACCOUNTS, \OPERATIONS]  \| [\PENDING, \ACCEPTED,  \MANAGERS, \CONTRACTORS, \TIME]}

@@

    \NodeTrans\\ [\epsilon[\FAIL] ::\EXPRS, \ACCOUNTS, \OPERATIONS]  \| [\PENDING, \ACCEPTED,  \MANAGERS, \CONTRACTORS, \TIME]}

@@

    \NEG\ \CHECKEXIST (\ADDR, \PENDING, \ACCEPTED)

>>>>

\begin{definition}%[Program]
\emph{Expressions} are defined by the following grammar:
\begin{align*}
  \CONSTANT & ::= \ADDR \mid \OPH \mid \PUH \mid\PUK \mid \CODE \mid
              \NTEZ \mid \SUNIT \mid \FALSE \mid \TRUE \\
  \STATUS & ::= \STATUSPENDING \mid \STATUSINCLUDING (\dots) \mid
            \STATUSTIMEOUT \\
  \ERROR &::= \ERRPRG \mid \ERRBAL \mid \ERRCOUNT \mid \ERRFEE \mid
           \ERRPUK \mid \ERRPUH \mid \ERRARG\\
	\EXPR & :: =  \CONSTANT \mid \STATUS \mid \ERROR \mid \VARIABLE \mid \lambda \VARIABLE. \EXPR\ 
	\mid \EXPR\EXPR \mid (\EXPR,\EXPR) \mid \FST\ \EXPR \mid \SND\ \EXPR 	%\mid \NI \EI
	\\&\mid \EXPR\ \PLUS\ \EXPR\ \mid \EXPR\ \MINUS\ \EXPR\ \mid \EXPR\ \EQUAL\ \EXPR\ \mid \EXPR\ \LESS\ \EXPR
	\\&\mid \EXPR\ \AND\ \EXPR\ \mid \EXPR\ \OR\ \EXPR\ \mid \NOT\ \EXPR
  \\&\mid \NIL\ \mid \CONS\ \EXPR\ \EXPR \mid \LEFT\ \EXPR \mid\RIGHT\
  \EXPR \mid \SOME\ \EXPR \mid \NONE
  \\& \mid \RAISE\ \EXPR \mid \TRY\ \EXPR\ \EXCEPT\ \EXPR
  \\& \mid \MATCH\ \EXPR\ \WITH\ \PATTERN\to\EXPR \dots
  \\& \mid \TRANSFER[\EXPR]\EXPR\EXPR\EXPR\EXPR
  \\& \mid \ORIGINATE\EXPR\EXPR\EXPR\EXPR\EXPR
  \\&\mid \QOP\  \EXPR\\
  \QOP & ::= balance \mid status \mid storage \mid code \mid counter
\end{align*}
\end{definition}

@@

A \emph{pending operation} is a tuple
$\langle  \OP, \OPH, \TIME, \STATUSPENDING\rangle $,
where $\OP$ is an operation, $\OPH$ is the operation hash,
and $\TIME$ is the time when the operation was injected.
Once it is accepted, it changes to an \emph{accepted operation}:
$\langle  \OP, \OPH, \TIME, \STATUSINCLUDED\ \TIME'\rangle $, where
$\TIME'$ is the  when it was included in the blockchain.

\begin{definition}%[Blockchain system]
A \emph{blockchain configuration} is a pair
$ \NODE \| \BLOCKCHAIN$ where
\fxnote[inline]{Should the model contain multiple nodes?}
\begin{enumerate}
\item $\NODE = [\ACCOUNTS, \OPERATIONS]$ is the state of a node, and
\item $\BLOCKCHAIN = [\PENDING, \ACCEPTED, \MANAGERS, \CONTRACTORS,
  \TIME]$ is the state of a blockchain  
  such that
  \begin{itemize}
  \item
    $\forall c \in \ACCOUNTS \implies \exists k\in \MANAGERS, k.\PUK =
    c.\PUK$;
  \item 
  \end{itemize}
\end{enumerate}
\end{definition}

@@

\begin{itemize}
\item library call \texttt{call\_contract} with \texttt{confirmations:None} corresponds to
  \begin{enumerate}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Inject'}
  \item{} first \TirName{Node-Op'}, \TirName{Node-Reject(balance,count,pubkey,fee)}
  \end{enumerate}
\item afterwards must check the status to find whether the
  transfer is still pending or \TirName{Block-Accept''} or
  \TirName{Block-Timeout''} has happened.
\item no further transfer is possible before the status is resolved to
  accept or timeout.
\end{itemize}

@@

    \NodeTrans\\ [\epsilon[\RAISE\ \ERRPUK] ::\EXPRS, \ACCOUNTS, \OPERATIONS]  \| [\PENDING, \ACCEPTED,  \MANAGERS, \CONTRACTORS, \TIME]}

@@

    \NodeTrans\\ [\epsilon[\RAISE\ \ERRPUH] ::\EXPRS, \ACCOUNTS, \OPERATIONS]  \| [\PENDING, \ACCEPTED,  \MANAGERS, \CONTRACTORS, \TIME]}

@@

    \NEG\ \CHECKEXIST (\OPH, \PENDING, \ACCEPTED)
